---
title: "Assignment3"
author: "Ankith Dasu"
date: "10/18/2022"
---

1) Formulating Transportation using R
```{r}
library(lpSolveAPI)
library(tinytex)
library(lpSolve)

```

#The data here is to be converted into a table format
```{r}
#cost matrix
cost_1 <- matrix(c(22,14,30,600,100,
                 16,20,24,625,120,
                 80,60,70,"-","-","210/220"),ncol = 5,nrow = 3,byrow = TRUE)
colnames(cost_1) <- c("Warehouse_1", "Warehouse_2","Warehouse_3","ProductionCost","Production Capacity")

rownames(cost_1) <- c("Plant_A","Plant_B","Demand")

cost_1
```
The Objective function is to Minimize the TC
$$  \text{Min} \hspace{.2cm} TC = 622 x_{11} + 614 x_{12} + 630 x_{13} + 641 x_{21} + 645 x_{22} + 649 x_{23}$$
Subject to the following constraints : Supply
$$X_{11} + X_{12} + X_{13}  >= 100$$
$$X_{21} + X_{22} + X_{23} >= 120$$
   
   
Subject to the following constraints : Demand
$$ X_{11} + X_{21} >= 80$$
$$ X_{12} + X_{22} >= 60$$
$$ X_{13} + X_{23} >= 70$$


Non-Negativity Constraints
$$ X_{ij} >= 0$$ 
Where i = 1,2 and j= 1,2,3
 
```{r}
#The capacity = 220 and Demand = 210. 
#A "Dummy" row for Warehouse_4 Is added here
trans.cost_1 <- matrix(c(622,614,630,0,100,
                       641,645,649,0,120,
                       80,60,70,10,220), ncol = 5, nrow = 3, byrow = TRUE)
trans.cost_1
```

```{r}
#Defining names for the rows and columns
colnames(trans.cost_1) <- c("Warehouse_1","Warehouse_2","Warehouse_3","Dummy","Production Capacity")

rownames(trans.cost_1) <- c("Plant_1", "Plant_2","Monthly Demand")
trans.cost_1
```
 
```{r}
#costs matrix
costs <- matrix(c(622,614,630,0,
                  641,645,649,0), nrow = 2, byrow = TRUE)
costs

```

We have to know that Supply function cannot be greater than the specified units where as the Demand function can be greater than the specified units.

```{r}
#setting up constraint signs and right-hand sides(supply side)
row.signs <- rep("<=",2)
row.rhs <- c(100,120)

#Demand side constraints#
col.signs <- rep(">=",4)
col.rhs <- c(80,60,70,10)

```

```{r}

lptrans <- lp.transport(costs, "min", row.signs,row.rhs,col.signs,col.rhs)

```

```{r}
lptrans$solution
```
80 AEDs in Plant 2 - Warehouse_1

60 AEDs in Plant 1 - Warehouse_2

40 AEDs in Plant 1 - Warehouse_3

30 AEDs in Plant 2 - Warehouse_3

The above should be the production in each plant and distribution to the three wholesaler warehouses to minimize the overall cost of production as well as shipping.

```{r}
lptrans$objval
```
The combined cost of production and shipping for the defibrilators is $132,790
```{r}
lptrans$duals
```


2)Formulate the dual of the transportation problem

Since the primal was to minimize the transportation cost the dual of it would be to maximize the value added(VA). 

U'v are the variables here.

```{r}
cost_2 <- matrix(c(622,614,630,100,"u1",
                   641,645,649,120,"u2",
                   80,60,70,220,"-",
                   "v1","v2","v3","-","-"),ncol = 5,nrow = 4,byrow = TRUE)
colnames(cost_2) <- c("Warehouse_1", "Warehouse_2","Warehouse_3","Production Capacity","Supply(Dual)")

rownames(cost_2) <- c("Plant_A","Plant_B","Demand","Demand(Dual)")
```




$$  \text{Max} \hspace{.2cm} VA = 100 P_{1} + 120 P_{2} +80 W_{1} + 60 W_{2} + 70 W_{3}$$

Total Profit Constraints
$$ u_{1} + v_{1} <= 622$$
$$ u_{1} + v_{2} <= 614$$
$$ u_{1} + v_{3} <= 630$$
$$ u_{2} + v_{1} <= 641$$
$$ u_{2} + v_{2} <= 645$$
$$ u_{2} + v_{3} <= 649$$
These are taken from the transposed matrix of the Primal of the LP. These are unrestricted where $$ u_{k} , v_{l}$$ where u=1,2 and v=1,2,3

```{r}
#Objective function

f.obj <- c(100,120,80,60,70)

#transposed from the constraints matrix in the primal
f.con <- matrix(c(1,0,1,0,0,
                  1,0,0,1,0,
                  1,0,0,0,1,
                  0,1,1,0,0,
                  0,1,0,1,0,
                  0,1,0,0,1), nrow = 6, byrow = TRUE)

f.dir <- c("<=",
           "<=",
           "<=",
           "<=",
           "<=",
           "<=")

f.rhs <- c(622,614,630,641,645,649)
lp("max",f.obj,f.con,f.dir,f.rhs)

lp("max",f.obj,f.con,f.dir,f.rhs)$solution
```
Z=139,120 and variables are:
$$ u_{1} = 614$$
$$ u_{2} = 633$$ 
$$ v_{1} = 8$$
$$ v_{3} = 16$$
So Z = $139,120 and variables are $$ u_{1} = 614$$ which represents Plant A
$$u_{2}=633$$ which represents Plant B
$$v_{1}=8$$ which represents Warehouse_1
$$v_{2}=16$$ which represents Warehouse_3


3) Economic Interpretation of the dua:

  minimal Z(Primal) = 132790 
  maximum Z(Dual) = 139120. 
  We understood that we should not be shipping from Plant(A/B) to all the three Warehouses. 
  Shipping should be done from

$$ 60 X_{12}$$ which is 60 Units from Plant A to Warehouse 2.
$$ 40 X_{13}$$ which is 40 Units from Plant A to Warehouse 3.
$$ 80 X_{13}$$ which is 60 Units from Plant B to Warehouse 1.
$$ 30 X_{13}$$ which is 60 Units from Plant B to Warehouse 3.

We will Max the profit from each distribution to the respective capacity.


$$ \hspace{2cm} u_{1} ^0 -v_{1} ^0 \le 622 $$ then we subtract $$ \hspace{2cm} v_{1} ^0 $$ to the other side to get  $$\hspace{2cm} u_{1} ^0\le 622 - v_{1} ^0 $$
To compute it would be $614 <= (-8+622) which is correct. we would continue to evaluate these equations:
$$ \hspace{2cm} u_{1}\le 622 - v_{1} => 614\le622 - 8 = 614=>correct$$
$$ \hspace{2cm} u_{1}\le 614 - v_{2} => 614\le614 - 0 = 614=>correct$$
$$ \hspace{2cm} u_{1}\le 630 - v_{3} => 614\le630 - 16 = 614=>correct$$
$$ \hspace{2cm} u_{2}\le 641 - v_{1} => 633\le614 - 8 = 633=>correct$$
$$ \hspace{2cm} u_{2}\le 645 - v_{2} => 633\le645 - 0 = 645=> Incorrect$$
$$ \hspace{2cm} u_{2}\le 649 - v_{3} => 633\le649 - 16 = 633=>correct$$
From the above using  the Sensitivity and Duality, the shadow price can be tested.
Change 100 to 101 and 120 to 121 in our LP Transport.

```{r}
row.rhs1 <- c(101,120)
row.signs1 <- rep("<=",2)
col.rhs1 <- c(80,60,70,10)
col.signs1 <- rep(">=",4)
row.rhs2 <- c(100,121)
row.signs2 <- rep("<=",2)
col.rhs2 <- c(80,60,70,10)
col.signs2 <- rep(">=",4)

lp.transport(costs,"min",row.signs,row.rhs,col.signs,col.rhs)

lp.transport(costs,"min",row.signs1,row.rhs1,col.signs1,col.rhs1)

lp.transport(costs,"min",row.signs2,row.rhs2,col.signs2,col.rhs2)

```

The 'min' of the specific function is taken here and we observe that the number goes down by 19. 
This indicates the shadow price is 19, that was found from the primal and adding 1 to each of the Plants.
Here, Plant B does not have a shadow price. 

From the dual variable $$ v_{1}$$ where Marginal Revenue <= Marginal Cost. The equation was $$ \hspace{2cm} u_{2}\le 645 - v_{2} => 633\le645 - 0 = 645=> Incorrect$$ and this was found by using $$ \hspace{2cm} u_{1} ^0 -v_{1} ^0 \le 622 $$ then we subtract $$ \hspace{2cm} v_{1} ^0 $$ to the other side to get  $$\hspace{2cm} u_{1} ^0\le 622 - v_{1} ^0 $$


The economic interpretation of the dual follows the universal rule of profit maximization which is MR >= MC 


```{r}
lp("max", f.obj,f.con, f.dir,f.rhs)$solution
```

Warehouse1 >= Plant1 + 621 i.e. MR1 >= MC1

Marginal Revenue which is the revenue generated for each additional unit sold relative to Marginal Cost (MC).
This is the change in cost at Plant 1 by inducing an increase in the supply function should be greater than or equal to the revenue generated for each additional unit distributed to Warehouse 1.

$$ 60 X_{12}$$ which is 60 Units from Plant A to Warehouse 2.
$$ 40 X_{13}$$ which is 40 Units from Plant A to Warehouse 3.
$$ 80 X_{13}$$ which is 60 Units from Plant B to Warehouse 1.
$$ 30 X_{13}$$ which is 60 Units from Plant B to Warehouse 3.

Out of 6, 5 of them had MR <= MC. 
Here we need MR=MC.
Plant B to Warehouse_2 does not satisfy the requirement. Hence, there will not be any AED device shipment.

