---
title: "Assignment6"
author: "Ankith Dasu"
date: "11/19/2022"
output: pdf_document
---



```{r}
library(lpSolveAPI)
```

```{r}
ap <- read.lp("ap.lp")
print(ap)
```

```{r}

Workers <- matrix(c("Sunday","Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday",18,27,22,26,25,21,19), ncol = 2, byrow = F)

colnames(Workers) <- c("Day", "Workers")

as.table(Workers)
```
Package handlers at AP are guaranteed a five-day work week with two consecutive days off. The base wage for the handlers is $750 per week. Workers working on Saturday or Sunday receive an additional $25 per day. 

```{r}
Days_off_and_wages <- matrix(c(1,2,3,4,5,6,7,
                               "Sunday and Monday","Monday and Tuesday", "Tuesday and Wednesday", "Wednesday and Thursday", "Thursday and Friday", "Friday and Saturday", "Saturday and Sunday","$775", "$800","$800", "$800", "$800", "$775", "$750"), ncol = 3, byrow = F)

colnames(Days_off_and_wages) <- c("Shift", "Days_off", "Wage")

as.table(Days_off_and_wages)
                               
```
```{r}

solve(ap)

```

```{r}
#Objective function (Total cost)

get.objective(ap)
```
Total cost to the firm after keeping total wage expense as low as possible while ensuring that there are sufficient workers available each day is $25,675


```{r}
#Variables (No of workers available each day)

get.variables(ap)
```

Findings:

No.of workers in shift1 = 2
No.of workers in shift2 = 4
No.of workers in shift3 = 5
No.of workers in shift4 = 0
No.of workers in shift5 = 8
No.of workers in shift6 = 1
No.of workers in shift7 = 13

Here, Workers available to work each day in terms of the objective function and constraints:

Workers on Sunday: x2+x3+x4+x5+x6 >= 18;
Workers on Monday: x3+x4+x5+x6+x7 >= 27;
Workers on Tuesday: x4+x5+x6+x7+x1 >= 22;
Workers on Wednesday: x5+x6+x7+x1+x2 >= 26;
Workers on Thursday: x6+x7+x1+x2+x3 >= 25;
Workers on Friday: x7+x1+x2+x3+x4 >= 21;
Workers on Saturday: x1+x2+x3+x4+x5 >= 19;
